Devito for large scale elastic modelling and anisotropic inversion

### 2020-04-20 paper.tex
typos-george

% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{9} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{9}  % for normal

@@

% Python style for highlighting
\newcommand\pythonstyle{\lstset{
    language=Python,
    basicstyle=\ttm,
    aboveskip=10pt,
    belowskip=10pt,
    otherkeywords={self},             % Add keywords here
    keywordstyle=\ttb\color{deepblue},
    emph={MyClass,__init__},          % Custom highlighting
    emphstyle=\ttb\color{deepred},    % Custom highlighting style
    stringstyle=\color{deepgreen},
    frame=tb,                         % Any extra options here
    showstringspaces=false            % 
}}

@@

\author{
\IEEEauthorblockN{1\textsuperscript{st} Louboutin Mathias}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Antlanta, GA \\
mlouboutin3@gatech.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
\IEEEauthorblockA{\textit{Devito Codes}}
\and
\IEEEauthorblockN{3\textsuperscript{th} Rhodri Nelson}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{4\textsuperscript{rd} Philipp Witte}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
\IEEEauthorblockA{\textit{TU-Delft}}
\and
\IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{8\textsuperscript{th} Geragd J. Gorman}
\IEEEauthorblockA{\textit{Imperial College London}}
}

@@

It is generally accepted that the PDE solver utilized within an
inversion workflow must satisfy the following criteria
\cite{virieuxmodelling}: - Efficient for multiple-source modeling - The
memory requirement of the modeling - The ability of a parallel algorithm
to use an increasing number of processors - Ability of the method to
process models of arbitrary levels of heterogeneity - Reduce the
nonlinearity of FWI - Feasibility of the extension of the modeling
approach to more realistic physical descriptions of the media.

@@

Previous work focused on the DSL and compiler to highlight the potential
application and use cases of Devito. Here, we present a series of
extensions and applications to large-scale three-dimensional problem
sizes as encountered in exploration geophysics. These experiments are
carried out in Cloud-based HPC systems and include elastic forward
modelling using distributed-memory parallelism and imaging based on the
tilted transverse isotropic (TTI) wave-equation
(\cite{virieux, thomsen1986, zhang-tti, duveneck, louboutin2018segeow}).
These proof of concepts highlight two critical features: first, the
ability of the symbolic interface and the compiler to translate to
large-scale adjoint-based inversion problems that require massive
compute (since thousands of PDEs are solved) as well as large amounts of
memory (since the adjoint state method requires the forward model to be
saved in memory). Secondly, through the elastic modelling example, we
demonstrate that \href{https://github.com/devitocodes/devito}{Devito}
now fully supports and automates vectorial and second order tensorial
staggered-grid finite-differences with the same high-level API
previously presented for a scalar fields defined on cartesian grids.

@@

\begin{python}
from devito import SparseTimeFunction
src = SparseTimeFunction(name="src", grid=grid,
                         npoint=1,
                         coordinates=x_s)
d = SparseTimeFunction(name="d", grid=grid,
                       npoint=1, nt=nt,
                       coordinates=x_r)
\end{python}

@@

\begin{python}
from devito import solve, Eq, Operator
eq = m * u.dt2 - u.laplace
update = Eq(u.forward, solve(eq, u.forward))
src_eqns = s1.inject(u.forward, expr=s1*dt**2/m)
d_eqns = d.interpolate(u)
\end{python}

@@

We here provide a succinct description of distributed-memory parallelism
in Devito; the interested reader should refer to the mpi tutorial at
\href{https://github.com/devitocodes/devito/blob/master/examples/mpi/overview.ipynb}{mpi-notebook}
for thorough explanations and practical examples.

@@

One of the main applications of seismic finite-difference modeling in
exploration geophysics is reverse-time migration (RTM), a wave-equation
based seismic imaging technique. Real-world seismic imaging presents a
number of challenges that make applying this method to industry-scale
problem sizes difficult. First of all, RTM requires an accurate
representation of the physics through sophisticated wave-equations such
as the tilted-transverse isotropic (TTI) wave-equation, for which both
forward and adjoint implementations have to be provided. Secondly,
wave-equations must be solved for a large number of independent
experiments, where each individual PDE solve in itself is expensive in
terms of FLOPs and memory usage. For certain workloads, limited domain
decomposition, that balances the domain size and the number of
independent experiment, and checkpointing techniques have to be applied.
In the following sections, we describe an industry-scale seismic imaging
problem presenting all of the aforementioned challenges, its
implementation with Devito, and the results of an experimentation
carried out on the Azure Cloud using a synthetic data set.

@@

In our seismic imaging case study, we use an anisotropic representation
of the physics called Tilted Transverse Isotropic (TTI) modeling
\cite{thomsen1986}. This representation for wave motion is one of the
most widely used in exploration geophysics since it captures the leading
order kinematics and dynamics of acoustic wave motion in highly
heterogeneous elastic media where the medium properties vary more
rapidly in the direction perpendicular to sedimentary strata
\cite{alkhalifah2000, baysal1983, bubetti2012, bubetti2014, bubesatti2016, chu2011, duveneck, fletcher, fowlertti2010, louboutin2018segeow, whitmore1983, witte2016segpve, xu2014, zhang2005, zhang2011, zhan2013}.
The TTI wave-equation is an acoustic, low dimensional (4 parameters, 2
wavefields) simplification of the 21 parameter and 12 wavefields
tensorial equations of motions \cite{hooke}. This simplified
representation is parametrized by the Thomsen parameters
$\epsilon(x), \delta(x)$ that relate to the global (many wavelength
propagation) difference in propagation speed in the vertical and
horizontal directions, and the tilt and azimuth angles
$\theta(x), \phi(x)$ that define the rotation of the vertical and
horizontal axis around the cartesian directions. However, unlike the
scalar isotropic acoustic wave-equation itself, the TTI wave-equation is
extremely computationally costly to solve and it is also not
self-adjoint as shown in \cite{louboutin2018segeow}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dx_u = cos(theta) * cos(phi) * u.dx + cos(theta) * sin(phi) * u.dy - sin(theta) * u.dz}
\NormalTok{dxx_u = (cos(theta) * cos(phi) * dx_u).dx.T + (cos(theta) * sin(phi) * dx_u).dy.T - (sin(theta) * dx_u).dz.T}
\end{Highlighting}
\end{Shaded}

@@

\begin{table}
\centering
\begin{tabular}{lll}
\toprule\addlinespace
spatial order & w/o optimizations & w/ optimizations\tabularnewline
\midrule
4 & 501 & 95\tabularnewline
8 & 539 & 102\tabularnewline
12 & 1613 & 160\tabularnewline
16 & 5489 & 276\tabularnewline
\bottomrule
\end{tabular}
\caption{Per-grid-point FLOPs of the finite-difference stencil for the
TTI wave-equation with different spatial discretization
orders.}\label{ttiFLOPs}
\end{table}

@@

With such complex physics, mathematics, and engineering, it becomes
evident that the implementation of a solver for this wave-equation is
exceptionally time-consuming and can lead to thousands of lines of code
even for a single type of discretization. The verification of the result
is no less complicated, since any small error is effectively untrackable
and any change to the finite-difference scheme or to the time-stepper is
difficult to achieve without substantial re-coding. Another complication
stems from the fact that practitioners of seismic inversion are often
geoscientists and not computer scientists/programmers. Unfortunately,
this background often either results in poorly written low performant
codes or it leads to complications when research codes are handed off to
computer scientists who know how to write fast codes but who often miss
the necessary geophysical domain knowledge. Neither situation is
conducive to addressing the complexities that come with implementing
codes based on the latest geophysical insights and high-performance
computing. With \href{https://github.com/devitocodes/devito}{Devito} on
the other hand, both the forward and adjoint equations can be
implemented in a few lines of Python code as illustrated with the
rotated operator in \#rotxpy.

@@

We now demonstrate the scalability of
\href{https://github.com/devitocodes/devito}{Devito} to real-world
applications with the imaging of an industry-scale three-dimensional TTI
subsurface model. This imaging was carried out in the Cloud on Azure and
takes advantage of recent work to port conventional cluster code to the
Cloud using a serverless approach. The serverless implementation is
described in detail in \cite{witte2019TPDedas, witte2019SEGedw} and
describes the steps to run computationnaly and financially efficient HPC
workloads in the Cloud. This imaging project, in collaboration with
Azure, demonstrates the scalability and robustness of
\href{https://github.com/devitocodes/devito}{Devito} to large scale
wave-equation based inverse problems and its cost-effectiveness in
combination with a serverless implementation of seismic imaging in the
Cloud. In this example, we imaged a synthetic three-dimensional
anisotropic subsurface model that mimics a realistic industry size
problem with a realistic representation of the physics (TTI). The
physical size of the problem is \texttt{10kmx10kmx2.8km} discretized on
a \texttt{12.5m} grid with 40 points of absorbing layer on each sides
that leads to \texttt{881x881x371} computational grid points (300
Million grid points). The final image is the sum of 1500 single-source
images and we computed 100 single-source images in parallel on 200 nodes
available using two ndes per source experiment.

@@

We also observe that if we had had access to more resources, and because
the imaging problem is embarassingly parallel over sources and can scale
to any computational ressource, we could have attempted imaging of all
of the 1500 sources in parallel, which theoretically leads to a
performance of 0.4PFLOP/s.

@@

\cite{witte2019TPDedas} fully describes the serverless implementation
of seismic inverse problems, including iterative algorithms for
least-square minimization problems (LSRTM). The 3D anisotropic imaging
results were presented as part of a keynote presentation at the EAGE HPC
workshop in October 2019 \cite{herrmann2019EAGEHPCaii} and at the Rice
O\&G HPC workshop \cite{witte2019RHPCssi} where we focused on the
serverless implementation of seismic inversion algorithms in the Cloud.
This work illustrates the flexibility and portability of Devito, as we
were able to easily port a code only tested and developed on local
hardware to the Cloud, with only minor adjustments. This portability
included the possibility to run MPI-based code for domain decomposition
in the Cloud, after developing it on a desktop computer. Our experiments
are reproducible using the instructions in a public repository
\href{https://github.com/slimgroup/Azure2019}{AzureTTI}, which contains,
among the other things, the Dockerfiles and Azure
\href{https://batch-shipyard.readthedocs.io}{batch-shipyard} setup. This
example can also easily ber un on a traditional HPC cluster environment
using for exaple \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or
Dask \cite{dask} for parallelization over sources.

@@

While the subsurface image in the previous Section is obtained with
anisotropic propagators capable of mimicing the real physics, in order
to model both the kinematics and the amplitudes correctly elastic
propagators are required. These propagators are for example extremely
important for global seismology, as the Shear surface waves (component
ignored in TTI) are the most hazardous ones. In this section, we exploit
the tensor algebra language introduced in
\href{https://github.com/devitocodes/devito}{Devito} v4.0 to express
elastic waves in a compact and elegant notation.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{v = VectorTimeFunction(name=}\StringTok{'v'}\NormalTok{, grid=model.grid, space_order=so, time_order=}\DecValTok{1}\NormalTok{)}
\NormalTok{tau = TensorTimeFunction(name=}\StringTok{'t'}\NormalTok{, grid=model.grid, space_order=so, time_order=}\DecValTok{1}\NormalTok{)}

\NormalTok{u_v = Eq(v.forward, model.damp * (v + s/rho*div(tau)))}
\NormalTok{u_t = Eq(tau.forward,  model.damp *  (tau + s * (l * diag(div(v.forward)) + mu * (grad(v.forward) + grad(v.forward).T))))}
\end{Highlighting}
\end{Shaded}

@@

We show the elastic particle velocity and stress for a broadly
recognized 2D synthetic model, the elastic
Marmousi-ii\cite{versteeg927, marmouelas} model. The wavefields are
shown on Figure~\ref{ElasWf} and its corresponding elastic shot records
are displayed in Figure~\ref{ElasShot}. Thes two figures show that the
wavefield is as expected purely acoustic in the water layer
(\texttt{t\_\{xy\}=0}) and transition perfectly at the ocean bottom to
an elastic wavefield. We can also clearly see the Shear wave wavefront
in the subsurface (around 1km depth). Thes two plots show that we
effectively solve the elastic wave-equation with an high-level
mathematical definition of the PDE that abstracts away all the
computational technicalities such as staggered grid analysis. The shot
records in Figure~\ref{ElasShot} also match the original data generated
by the creator of the elastic model in \cite{marmouelas}.

@@

These 21 fields, with the grid we just describe, leads to a minimum of
461Gb of memory for modelling only. For this experiment, we obtained
access to small HPC VM on azure called \texttt{Standard\_H16r} that are
16 cores Intel Xeon E5 2667 v3 with no hyperthreading and used 32 nodes
for a single source experiment (we solved a single wave-equation). We
used a 12th order discretization that leads to 2.8TFLOP/time-step to be
computed for this model and propagated the elastic wave for 16 seconds
(23000 time steps). The modelling ran in 16 hours which converts to
1.1TFLOP/s. While these number may seem low, the elastic kernel is
extremely memory bound, while the TTI kernel is nearly compute bound
(see rooflines in \cite{louboutin2016ppf, devito-api, devito-compiler})
making it more computationally efficient, in particular in combination
with MPI. Future work with involve working on the InfiniBand enabled and
true HPC VM on azure to achieve Cloud performance on par with state of
the art Cluster performance. From the performance obtained in this
experiement, and assuming a fairly standard setup of 5000 independent
source experiements, computing a synthetic dataset would require 322
EFLOP to be computed (23k time-steps x 2.8TFLOP/time-step x 500
sources), or using the full scalability and computing all sources in
parallel would lead to 5.5PFLOP/s from our single source experiement
result.

@@

Because earlier performance benchmark focuse mainly on roofline model
analysis, we also compared the runtime of
\href{https://github.com/devitocodes/devito}{Devito} with a reference
open source hand-coded propagator in collaboration with its author for
compeleteness. This propagator, described in \cite{thorbecke} is a
state of the art elastic kernel (Equation~\ref{elas1}). For a fair
comparison, we ensure, in collaboration with the author of
\href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} that the
physical and computational settings were identical. The setting were the
following:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  2001 by 1001 physical grid points.
\item
  200 grid points of dampening layer (absorbing layer \cite{cerjan}) on
  all four sides (total of 2401x1401 computational grid points).
\item
  10001 time steps.
\item
  Single point source, 2001 receivers.
\item
  Same compiler (gcc/icc) to compile
  \href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} and
  run Devito.
\item
  Intel(R) Xeon(R) CPU E3-1270 v6 @ 3.8GHz.
\item
  Single socket, four physical cores, four physical threads, thread
  pinning to cores and hyperthreading off.
\end{itemize}

>>>>

% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8.5} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8.5}  % for normal

@@

% Python style for highlighting
\newcommand\pythonstyle{\lstset{
    language=Python,
    basicstyle=\ttm,
    aboveskip=10pt,
    belowskip=10pt,
    otherkeywords={self},             % Add keywords here
    keywordstyle=\ttb\color{deepblue},
    emph={MyClass,__init__},          % Custom highlighting
    emphstyle=\ttb\color{deepred},    % Custom highlighting style
    stringstyle=\color{deepgreen},
    numbers=none,
    frame=tb,                         % Any extra options here
    showstringspaces=false            % 
}}

@@

\author{
\IEEEauthorblockN{1\textsuperscript{st} Louboutin Mathias}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Antlanta, GA \\
mlouboutin3@gatech.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
\IEEEauthorblockA{\textit{Devito Codes}}
\and
\IEEEauthorblockN{3\textsuperscript{th} Rhodri Nelson}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{4\textsuperscript{rd} Philipp Witte}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
\IEEEauthorblockA{\textit{TU-Delft}}
\and
\IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
\IEEEauthorblockA{\textit{Imperial College London}}
}

@@

It is generally accepted that the PDE solver utilized within an
inversion workflow must satisfy the following criteria
\cite{virieuxmodelling}: - Efficient for multiple-source modeling - The
memory requirement of the modeling - The ability of a parallel algorithm
to use an increasing number of processors - Ability of the method to
process models of arbitrary levels of heterogeneity - Reduce the
non-linearity of FWI - Feasibility of the extension of the modeling
approach to more realistic physical descriptions of the media.

@@

Previous work focused on the DSL and compiler to highlight the potential
application and use cases of Devito. Here, we present a series of
extensions and applications to large-scale three-dimensional problem
sizes as encountered in exploration geophysics. These experiments are
carried out in Cloud-based HPC systems and include elastic forward
modelling using distributed-memory parallelism and imaging based on the
tilted transverse isotropic (TTI) wave-equation
(\cite{virieux, thomsen1986, zhang-tti, duveneck, louboutin2018segeow}).
These proof of concepts highlight two critical features: first, the
ability of the symbolic interface and the compiler to translate to
large-scale adjoint-based inversion problems that require massive
compute (since thousands of PDEs are solved) as well as large amounts of
memory (since the adjoint state method requires the forward model to be
saved in memory). Secondly, through the elastic modelling example, we
demonstrate that \href{https://github.com/devitocodes/devito}{Devito}
now fully supports and automates vectorial and second order tensorial
staggered-grid finite-differences with the same high-level API
previously presented for scalar fields defined on cartesian grids.

@@

\begin{python}
from devito import SparseTimeFunction
s = SparseTimeFunction(name="src",
                       grid=grid, npoint=1,
                       coordinates=x_s)
d = SparseTimeFunction(name="d", grid=grid,
                       npoint=1, nt=nt,
                       coordinates=x_r)
\end{python}

@@

\begin{python}
from devito import solve, Eq, Operator
eq = m * u.dt2 - u.laplace
update = Eq(u.forward, solve(eq, u.forward))
src_eqns = s.inject(u.forward, expr=s*dt**2/m)
d_eqns = d.interpolate(u)
\end{python}

@@

We here provide a succinct description of distributed-memory parallelism
in Devito; the interested reader should refer to the MPI tutorial at
\href{https://github.com/devitocodes/devito/blob/master/examples/mpi/overview.ipynb}{mpi-notebook}
for thorough explanations and practical examples.

@@

One of the main applications of seismic finite-difference modeling in
exploration geophysics is reverse-time migration (RTM), a wave-equation
based seismic imaging technique. Real-world seismic imaging presents a
number of challenges that make applying this method to industry-scale
problem sizes difficult. First of all, RTM requires an accurate
representation of the physics through sophisticated wave-equations such
as the tilted-transverse isotropic (TTI) wave-equation, for which both
forward and adjoint implementations have to be provided. Secondly,
wave-equations must be solved for a large number of independent
experiments, where each individual PDE solve in itself is expensive in
terms of FLOPs and memory usage. For certain workloads, limited domain
decomposition, that balances the domain size and the number of
independent experiments, and checkpointing techniques have to be applied.
In the following sections, we describe an industry-scale seismic imaging
problem presenting all of the aforementioned challenges, its
implementation with Devito, and the results of an experimentation
carried out on the Azure Cloud using a synthetic data set.

@@

In our seismic imaging case study, we use an anisotropic representation
of the physics called Tilted Transverse Isotropic (TTI) modeling
\cite{thomsen1986}. This representation for wave motion is one of the
most widely used in exploration geophysics since it captures the leading
order kinematics and dynamics of acoustic wave motion in highly
heterogeneous elastic media where the medium properties vary more
rapidly in the direction perpendicular to sedimentary strata
\cite{alkhalifah2000, baysal1983, bubetti2012, bubetti2014, bubesatti2016, chu2011, duveneck, fletcher, fowlertti2010, louboutin2018segeow, whitmore1983, witte2016segpve, xu2014, zhang2005, zhang2011, zhan2013}.
The TTI wave-equation is an acoustic, low dimensional (4 parameters, 2
wavefields) simplification of the 21 parameters and 12 wavefields
tensorial equations of motions \cite{hooke}. This simplified
representation is parametrized by the Thomsen parameters
$\epsilon(x), \delta(x)$ that relate to the global (many wavelength
propagation) difference in propagation speed in the vertical and
horizontal directions, and the tilt and azimuth angles
$\theta(x), \phi(x)$ that define the rotation of the vertical and
horizontal axis around the cartesian directions. However, unlike the
scalar isotropic acoustic wave-equation itself, the TTI wave-equation is
extremely computationally costly to solve and it is also not
self-adjoint as shown in \cite{louboutin2018segeow}.

@@

\begin{python}
dx_u = cos(theta) * cos(phi) * u.dx +
       cos(theta) * sin(phi) * u.dy -
       sin(theta) * u.dz
dxx_u = (cos(theta) * cos(phi) * dx_u).dx.T +
        (cos(theta) * sin(phi) * dx_u).dy.T -
        (sin(theta) * dx_u).dz.T
\end{python}

@@

\begin{table}
\centering
\begin{tabular}{ccc}
\toprule\addlinespace
spatial order & w/o optimizations & w/ optimizations\tabularnewline
\midrule
4 & 501 & 95\tabularnewline
8 & 539 & 102\tabularnewline
12 & 1613 & 160\tabularnewline
16 & 5489 & 276\tabularnewline
\bottomrule
\end{tabular}
\caption{Per-grid-point FLOPs of the finite-difference stencil for the
TTI wave-equation with different spatial discretization
orders.}\label{ttiFLOPs}
\end{table}

@@

With such complex physics, mathematics, and engineering, it becomes
evident that the implementation of a solver for this wave-equation is
exceptionally time-consuming and can lead to thousands of lines of code
even for a single type of discretization. The verification of the result
is no less complicated, since any small error is effectively untrackable
and any change to the finite-difference scheme or to the time-stepper is
difficult to achieve without substantial re-coding. Another complication
stems from the fact that practitioners of seismic inversion are often
geoscientists and not computer scientists/programmers. Unfortunately,
this background often either results in poorly written low performant
codes or it leads to complications when research codes are handed off to
computer scientists who know how to write fast codes but who often miss
the necessary geophysical domain knowledge. Neither situation is
conducive to addressing the complexities that come with implementing
codes based on the latest geophysical insights and high-performance
computing. With \href{https://github.com/devitocodes/devito}{Devito} on
the other hand, both the forward and adjoint equations can be
implemented in a few lines of Python code as illustrated with the
rotational operator in \#rotxpy.

@@

We now demonstrate the scalability of
\href{https://github.com/devitocodes/devito}{Devito} to real-world
applications with the imaging of an industry-scale three-dimensional TTI
subsurface model. This imaging was carried out in the Cloud on Azure and
takes advantage of recent work to port conventional cluster code to the
Cloud using a serverless approach. The serverless implementation is
described in detail in \cite{witte2019TPDedas, witte2019SEGedw} and
describes the steps to run computationnaly and financially efficient HPC
workloads in the Cloud. This imaging project, in collaboration with
Azure, demonstrates the scalability and robustness of
\href{https://github.com/devitocodes/devito}{Devito} to large scale
wave-equation based inverse problems and its cost-effectiveness in
combination with a serverless implementation of seismic imaging in the
Cloud. In this example, we imaged a synthetic three-dimensional
anisotropic subsurface model that mimics a realistic industry size
problem with a realistic representation of the physics (TTI). The
physical size of the problem is \texttt{10kmx10kmx2.8km} discretized on
a \texttt{12.5m} grid with 40 points of absorbing layer on each sides
that leads to \texttt{881x881x371} computational grid points (300
Million grid points). The final image is the sum of 1500 single-source
images and we computed 100 single-source images in parallel on 200 nodes
available using two nodes per source experiment.

@@

We also observe that if we had had access to more resources, and because
the imaging problem is embarassingly parallel over sources and can scale
to any computational resources, we could have attempted imaging of all
of the 1500 sources in parallel, which theoretically leads to a
performance of 0.4PFLOP/s.

@@

\cite{witte2019TPDedas} fully describes the serverless implementation
of seismic inverse problems, including iterative algorithms for
least-square minimization problems (LSRTM). The 3D anisotropic imaging
results were presented as part of a keynote presentation at the EAGE HPC
workshop in October 2019 \cite{herrmann2019EAGEHPCaii} and at the Rice
O\&G HPC workshop \cite{witte2019RHPCssi} where we focused on the
serverless implementation of seismic inversion algorithms in the Cloud.
This work illustrates the flexibility and portability of Devito, as we
were able to easily port a code only tested and developed on local
hardware to the Cloud, with only minor adjustments. This portability
included the possibility to run MPI-based code for domain decomposition
in the Cloud, after developing it on a desktop computer. Our experiments
are reproducible using the instructions in a public repository
\href{https://github.com/slimgroup/Azure2019}{AzureTTI}, which contains,
among the other things, the Dockerfiles and Azure
\href{https://batch-shipyard.readthedocs.io}{batch-shipyard} setup. This
example can also easily ber un on a traditional HPC cluster environment
using for example \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or
Dask \cite{dask} for parallelization over sources.

@@

While the subsurface image in the previous Section is obtained with
anisotropic propagators capable of mimicking the real physics, in order
to model both the kinematics and the amplitudes correctly elastic
propagators are required. These propagators are for example extremely
important for global seismology, as the Shear surface waves (component
ignored in TTI) are the most hazardous ones. In this section, we exploit
the tensor algebra language introduced in
\href{https://github.com/devitocodes/devito}{Devito} v4.0 to express
elastic waves in a compact and elegant notation.

@@

\begin{python}
from devito import VectorTimeFunction,
                   TensorTimeFunction
v = VectorTimeFunction(name="v", grid=grid,
                       space_order=so,
                       time_order=1)
tau = TensorTimeFunction(name="t", grid=grid,
                         space_order=so,
                         time_order=1)
u_v = Eq(v.forward,
         damp * (v + s/rho*div(tau)))
u_t = Eq(tau.forward,
         damp * (tau + s * (l * diag(div(v.forward)) + mu * (grad(v.forward) + grad(v.forward).T))))
\end{python}

@@

We show the elastic particle velocity and stress for a broadly
recognized 2D synthetic model, the elastic
Marmousi-ii\cite{versteeg927, marmouelas} model. The wavefields are
shown on Figure~\ref{ElasWf} and its corresponding elastic shot records
are displayed in Figure~\ref{ElasShot}. Thes two figures show that the
wavefield is as expected purely acoustic in the water layer
(\texttt{t\_\{xy\}=0}) and transition perfectly at the ocean bottom to
an elastic wavefield. We can also clearly see the Shear wave wavefront
in the subsurface (around 1km depth). Thes two plots show that we
effectively solve the elastic wave-equation with a high-level
mathematical definition of the PDE that abstracts away all the
computational technicalities such as staggered grid analysis. The shot
records in Figure~\ref{ElasShot} also match the original data generated
by the creator of the elastic model in \cite{marmouelas}.

@@

These 21 fields, with the grid we just describe, leads to a minimum of
461Gb of memory for modelling only. For this experiment, we obtained
access to small HPC VM on azure called \texttt{Standard\_H16r} that are
16 cores Intel Xeon E5 2667 v3 with no hyperthreading and used 32 nodes
for a single source experiment (we solved a single wave-equation). We
used a 12th order discretization that leads to 2.8TFLOP/time-step to be
computed for this model and propagated the elastic wave for 16 seconds
(23000 time steps). The modelling ran in 16 hours which converts to
1.1TFLOP/s. While these number may seem low, the elastic kernel is
extremely memory bound, while the TTI kernel is nearly compute bound
(see rooflines in \cite{louboutin2016ppf, devito-api, devito-compiler})
making it more computationally efficient, in particular in combination
with MPI. Future work with involve working on the InfiniBand enabled and
true HPC VM on azure to achieve Cloud performance on par with state of
the art Cluster performance. From the performance obtained in this
experiment, and assuming a fairly standard setup of 5000 independent
source experiments, computing a synthetic dataset would require 322
EFLOP to be computed (23k time-steps x 2.8TFLOP/time-step x 500
sources), or using the full scalability and computing all sources in
parallel would lead to 5.5PFLOP/s from our single source experiment
result.

@@

Because earlier performance benchmarks focus mainly on roofline model
analysis, we also compared the runtime of
\href{https://github.com/devitocodes/devito}{Devito} with a reference
open source hand-coded propagator in collaboration with its author for
completeness. This propagator, described in \cite{thorbecke} is a
state of the art elastic kernel (Equation~\ref{elas1}). For a fair
comparison, we ensure, in collaboration with the author of
\href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} that the
physical and computational settings were identical. The setting were the
following:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  2001 by 1001 physical grid points.
\item
  200 grid points of dampening layer (absorbing layer \cite{cerjan}) on
  all four sides (total of 2401x1401 computational grid points).
\item
  10001 time steps.
\item
  Single point source, 2001 receivers.
\item
  Same compiler (GCC/ICC) to compile
  \href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} and
  run Devito.
\item
  Intel(R) Xeon(R) CPU E3-1270 v6 @ 3.8GHz.
\item
  Single socket, four physical cores, four physical threads, thread
  pinning to cores and hyperthreading off.
\end{itemize}
